Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2018-06-10 paper/referee_reply1.txt

4) There should be an uncertainty on w_p in the mock data, i.e. on the lines in Figs. 4-6. Since the model is probabilistic (both the abundance matching step and size assignment have random scatters), different Monte Carlo realisations of the model should give different results. Further, for a given realisation jackknife resampling would give an uncertainty estimate in a similar manner to the SDSS data. I would like to see these uncertainties as bands on the plots, or at least a general estimate of their magnitude. Looking at the bottom-right panel of Fig. 6, for the green line to be consistent with 0 the uncertainty on it must be fairly large, at least at low r; without this uncertainty it is hard to gauge the significance of the difference between the purple line and the data.

5) When some of the stellar mass is removed due to stripping, does this not cause the final SMF that's actually used in the model to disagree with the observed input SMF? In the discussion just below Eq 1, it seems that the distribution of M* is the one that matches the observations, not M'*. Or is the SMF that's input to AM modified somehow so that only after this stripping does it match Moustakas et al 2013?

Reply) The implementation of the stellar mass stripping model is indeed crude and to some extent the self-inconsistency pointed out by the referee is inescapable: we are after all not providing a model for the true evolution of post-infall stellar mass, but instead we are grafting an approximate treatment that is intended to approximate the impact of tidal stripping in a statistical sense. That being said, the modification to the SMF is rather small, at or below 10% for the mass ranges relevant to our clustering measurements, and so comparing these two models directly should not be too misleading. We have added a sentence at the end of 3.2.1 calling attention to this detail.

### 2018-06-10 paper/referee_reply1.txt

1) The size-halo connection that the authors advocate, presented in Section 3.2.2, sets each galaxy to be 1% of the size of its halo at the time of Mpeak. The authors state below Eq 3 that this defines a _median_ relation between Reff and Mhalo, around which they assume a log-normal scatter of 0.2 dex. However, it seems that a more direct application of Eq 2 would be to apply it on a halo-by-halo basis, since Rpeak is known for every object in the catalogue. This would specify the full Reff distribution, with scatter deriving from the scatter in Rpeak. Why is this not done? No justification is given for using a log-normal scatter of 0.2 dex.

Reply) The concluding sentence to Section 3.2.2 was indeed confusing; it turns out it was also misleading: in fact, in our analysis we assigned R_1/2 in exactly the fashion the referee described, about each halo's own R_Mpeak. We have changed the language to more accurately reflect the implementation of the model, and thank the referee for pointing out this confusing language.

2) This model is compared to one in which Reff does not correlate with any halo property at fixed stellar mass, but simply depends on M* in a manner that reproduces the M*-Reff relation by construction. Again the authors take 0.2 dex scatter around the median relation, but if one really wanted to reproduce the M*-Reff relation -- including its scatter -- one would take this scatter in bins of M* from the data. Even better would be to drop the assumption of Gaussian scatter altogether and simply assign a size randomly to each model galaxy by sampling the observed Reff distribution at given M*. It appears the authors' choice is motivated by a desire for compatibility with Somerville et al 2017, but it is unclear why this is especially relevant or important here.

Reply) We agree with the referee that drawing directly from P_SDSS(R_1/2 | M*) would improve the fidelity of our model with respect to the SDSS. However, our goal here is actually twofold. On the one hand, we want to identify a model that has sufficient flexibility to capture size-dependent clustering with good accuracy, using SDSS as our validation dataset. On the other hand, we are attempting to build a class of models that can be applied equally well to other datasets, including those may not have the statistics of the SDSS, as is the case in present-day high-redshift surveys. We have added what we hope is a clarifying paragraph on this point towards the end of Section 7.

@@

Reply) The implementation of the stellar mass stripping model is indeed crude and to some extent the self-inconsistency pointed out by the referee is inescapable: we are after all not providing a model for the true evolution of post-infall stellar mass, but instead we are grafting an approximate treatment that is intended to approximate the impact of tidal stripping in a statistical sense. That being said, the modification to the SMF is rather small, at or below 10% for the mass ranges relevant to our clustering measurements, and so comparing these two models directly should not be too misleading. We have added a sentence at the end of 3.2.1 calling attention to this detail.

6) A complementary method for studying the size-halo connection is to constrain correlations of galaxy size with halo properties *at fixed stellar mass*, which has been done using both weak lensing (e.g. Sonnenfeld et al 2015, Sonnenfeld & Leauthaud 2017) and galaxy dynamics (e.g. Desmond 2016, 2017). These are means of testing disk models such as Dutton et al 2007 & 2011 which the authors mention. It would be helpful to provide more information on how the Rvir-based model presented here relates to this alternative method. For example, since on average M* is positively correlated with Rpeak by AM, and empirically Reff is positively correlated with M*, then the model in which Reff depends only on M* (Sec 3.2.1) will also positively correlate Reff with Rpeak. It is not clear how this correlation differs from Reff=0.01*Rpeak, or in other words how the latter assumption correlates Reff with halo properties at fixed M*. It would be interesting to see 1) The Reff-Rpeak relation implied by the model in Sec 3.2.1, and/or 2) The correlation of Reff with halo mass or concentration at fixed M* implied by Reff=0.01*Rpeak. Understanding these different approaches will be of great help in devising a unified model for the size-halo connection.

Reply)  We have verified the correlation between R_1/2 and Mvir at fixed M* in present day centrals, and added a paragraph to Section 6 with the suggested references.

@@

7) Final paragraph of Section 6: It has been known for almost 20 years that the scatter in halo spin at fixed mass is larger than the scatter in galaxy size (de Jong & Lacey 2000, Gnedin et al 2007). The forward modelling implications of Reff^{disk} \propto \lambda have already been explored by Desmond & Wechsler 2015, who find it to be problematic for the Tully-Fisher relation as well as the scatter in size. Further, most modern hydrodynamical simulations (including Illustris [Rodriguez-Gomez et al 2017], EAGLE [Sales et al 2009, Desmond et al 2017] and Magneticum [Teklu et al 2015]) correlate Reff little if at all with halo spin. (It would be useful in general to provide a qualitative comparison with hydro sims, which have the potential to test models for the size-halo connection from first principles, and are beginning to do so.) Again it is unclear to me why Somerville et al 2017 should be of any particular importance for the scatter: The scatter in Reff at fixed M* comes straight from the data, and the scatter in lambda at fixed Mhalo (only relevant for angular momentum-based size models, which are barely mentioned here) comes straight from the N-body sim.

Reply) We have added a new sentence that clarifies the purpose of the final paragraph of Section 6, which is to call attention to the methodological similarity of our model to that in Somerville+17. We are interested in applying our forward-modeling techniques to see what insight can be learned about the R_1/2^{disk} \propto \lambda from galaxy clustering, which would be complementary to the knowledge gained by the Desmond & Wechsler 2015 study.

### 2018-06-10 paper/galsize_paper.tex
Modified reply to comment 5 and implemented revision in draft
Revised text and reply again to comment 6

It is natural to consider the possibility that stellar mass is stripped from satellite galaxies after infall. The basis for our implementation of this phenomenon is the fitting function presented in \citet{smith_etal16}, which was calibrated by studying stellar mass loss in a suite of high-resolution hydrodynamical simulations. In this model, $f_{\ast}$ quantifies the remaining fraction of stellar mass as a function of $f_{\rm DM},$ the fraction of dark matter mass that remains after subhalo infall:
\beq
f_{\ast} = 1 - \exp(-14.2f_{\rm DM}).
\eeq
For $f_{\rm DM}$ we use the ratio of present-day subhalo mass divided by the peak mass, $M_{\rm vir}/M_{\rm peak}.$ If we denote the post-stripping stellar mass as $M_{\ast}',$ then we have $M_{\ast}'\equiv f_{\ast}M_{\ast},$ where $M_{\ast}$ is given by the initial application of abundance matching. We then calculate the post-stripping radius by interpolating $\langle\rhalf'\vert\mstar'\rangle$ directly from SDSS data. In this way, we model satellite mass loss and  size truncation in a manner that mimics what is seen in hydrodynamical simulations. As shown in \citet{smith_etal16}, stellar stripping in this model does not begin until $f_{\rm DM}\lesssim0.2,$ and in order for $\sim50\%$ of stars to be stripped, the DM halo needs to be stripped to $\sim5\%$ of its initial mass.

@@

To date, the $\rhalf-$dependence of $\Delta\Sigma$ has not yet been measured for $\mstar-$complete samples. In the left panels of Figure \ref{fig:lensingprediction}, we show predictions of the $\rvir-$based model for future measurements of $\Delta\Sigma.$ The halo model principles underlying the lensing trends are the same as for $\wproj.$ For $\mstar\lesssim10^{11}\msun,$ the difference in $F_{\rm sat}$ between small and large samples is the dominant factor; satellites occupy higher mass host halos relative to centrals of the same stellar mass, boosting $\Delta\Sigma$ for small relative to large galaxy samples. At sufficiently large stellar mass $\mstar\gtrsim10^{11}\msun,$ when $F_{\rm sat}\lesssim10-15\%,$ the role of centrals becomes dominant; the $\rvir-$based model predicts that $\mhalo$ increases with $\rhalf$ for centrals, which is reflected in the reversal of the small-scale signal in the bottom left panel.

>>>>

It is natural to consider the possibility that stellar mass is stripped from satellite galaxies after infall. The basis for our implementation of this phenomenon is the fitting function presented in \citet{smith_etal16}, which was calibrated by studying stellar mass loss in a suite of high-resolution hydrodynamical simulations. In this model, $f_{\ast}$ quantifies the remaining fraction of stellar mass as a function of $f_{\rm DM},$ the fraction of dark matter mass that remains after subhalo infall:
\beq
f_{\ast} = 1 - \exp(-14.2f_{\rm DM}).
\eeq
For $f_{\rm DM}$ we use the ratio of present-day subhalo mass divided by the peak mass, $M_{\rm vir}/M_{\rm peak}.$ If we denote the post-stripping stellar mass as $M_{\ast}',$ then we have $M_{\ast}'\equiv f_{\ast}M_{\ast},$ where $M_{\ast}$ is given by the initial application of abundance matching. We then calculate the post-stripping radius by interpolating $\langle\rhalf'\vert\mstar'\rangle$ directly from SDSS data. In this way, we model satellite mass loss and  size truncation in a manner that mimics what is seen in hydrodynamical simulations. As shown in \citet{smith_etal16}, stellar stripping in this model does not begin until $f_{\rm DM}\lesssim0.2,$ and in order for $\sim50\%$ of stars to be stripped, the DM halo needs to be stripped to $\sim5\%$ of its initial mass. We note that mass loss slightly depletes the stellar mass function of this model relative to the \citet{moustakas_etal13}, but the differences are less than $10\%$ for stellar masses $\mstar\gtrsim10^{9.5}\msun.$

@@

In other closely related work, efforts to constrain correlations between galaxy size and various halo properties at fixed {\em stellar mass} have been carried out using both weak lensing \citep[e.g.,][]{sonnenfeld_leauthaud18} and galaxy dynamics \citep[e.g.,][]{desmond_wechsler15,desmond_wechsler17}. Our results complement these previous studies: forward-modeling galaxies directly into simulations allows us to test hypotheses of correlations at fixed {\em halo mass.} Calculating fixed-$\mstar$ correlations from a Monte Carlo realization of the model is straightforward, and in our fiducial model, in which $\mean{\rhalf}{\rmpeak}=0.01\rmpeak,$ for Milky Way mass centrals of the same $\mstar$ there exists a significant positive correlation between present day halo mass $\mvir$ and galaxy size $\rhalf.$ Alternative observational methods based on lensing and dynamics such as those mentioned above are a natural testbed for the conclusions drawn here based on two-point clustering. We now give an example of how such a future measurement could be made to directly and independently test these conclusions.

The $\rhalf-$dependence of $\Delta\Sigma$ has not yet been measured for $\mstar-$complete samples. In the left panels of Figure \ref{fig:lensingprediction}, we show predictions of the $\rvir-$based model for future measurements of $\Delta\Sigma.$ The halo model principles underlying the lensing trends are the same as for $\wproj.$ For $\mstar\lesssim10^{11}\msun,$ the difference in $F_{\rm sat}$ between small and large samples is the dominant factor; satellites occupy higher mass host halos relative to centrals of the same stellar mass, boosting $\Delta\Sigma$ for small relative to large galaxy samples. At sufficiently large stellar mass $\mstar\gtrsim10^{11}\msun,$ when $F_{\rm sat}\lesssim10-15\%,$ the role of centrals becomes dominant; the $\rvir-$based model predicts that $\mhalo$ increases with $\rhalf$ for centrals, which is reflected in the reversal of the small-scale signal in the bottom left panel.
